---
title: "ADA Final"
author: "Sophie Bair"
date: "4/8/2021"
output: pdf_document
---

Significant
* Land area
* Number of hospital beds per capita
* Number of serious crimes per capita
* Percent high school graduates
* Percent below poverty level
* Per capita income
* Personal income
* Percent unemployment

```{r}
data<-read.table("/Users/seb2244/Downloads/APPENC02.txt", sep="", header=FALSE)
UNI<-2244
set.seed(UNI)
index <-sample(c(1:440))
mydata<-data[index[1:250],]
head(mydata)
```

```{r}
write.csv(mydata, "/Users/seb2244/Downloads/ada_project_df.csv")
```

```{r}
ne <- mydata[mydata$V17 == 1, ]
nc <- mydata[mydata$V17 == 2, ]
s <- mydata[mydata$V17 == 3, ]
w <- mydata[mydata$V17 == 4, ]
nrow(ne)
nrow(nc)
nrow(s)
nrow(w)
```
```{r}
length(unique(mydata$V1)) # no duplicates
```

```{r}
subset = mydata[, c(4:7, 11:20)]
cor(subset)
```
```{r}
# import packages
library(car)
```

V4: Land Area
```{r}
# visualize data 
boxplot(log(V4)~V17, data = mydata)$stats # west has large land area
```
```{r}
# test assumptions 
hist(mydata$V4, breaks = 50) # not normal, right skewed
leveneTest(V4~factor(V17), data = mydata) # significantly unequal variance
# independence assumption is assumed to be true since other than counties which are bordering and may influence crime rates in the other, counties can be considered independent 
```
```{r}
kruskal.test(V4~factor(V17), data = mydata) # significant
```
V5: Total population
```{r}
# test assumptions 
hist(mydata$V5, breaks = 50) # not normal! 
leveneTest(V4~factor(V17), data = mydata) # not equal variance
```
```{r}
boxplot(V5~V17, data = mydata)$stats # ne has large population
```

```{r}
kruskal.test(V5~factor(V17), data = mydata) # pop not significant
```

V6: Percent of population aged 18-34
```{r}
# test assumptions 
hist(mydata$V6, breaks = 50) # fairly normal
leveneTest(V6~factor(V17), data = mydata) # not equal variance, but it's probably fine since anova is robust to some difference in variance as long as its not too severe. I did anova and k-w to be safe
var(s$V6)
var(ne$V6)
var(nc$V6)
var(w$V6)
```
```{r}
boxplot(V6~V17, data = mydata)$stats # pretty consistent percent young
```
```{r}
summary(aov(V6~factor(V17), data = mydata)) # younger pop not significant
```
```{r}
kruskal.test(V6~factor(V17), data = mydata) # again, not significant
```

V7: percent of population aged 65+
```{r}
# test assumptions 
hist(mydata$V7, breaks = 50) # fairly normal
leveneTest(V7~factor(V17), data = mydata) # huge imbalance in variance
var(s$V7)
var(ne$V7)
var(nc$V7)
var(w$V7)
```
```{r}
boxplot(V7~V17, data = mydata)$stats # slightly higher older pop in ne
```

```{r}
kruskal.test(V7~factor(V17), data = mydata) # pop not significant
```

V8: number of active physicians 
```{r}
# test assumptions 
mydata$doctors_per_cap = mydata$V8/mydata$V5 # scale by population
hist(mydata$doctors_per_cap, breaks = 50) # right skewed
leveneTest(doctors_per_cap~factor(V17), data = mydata) # equal var
```
```{r}
boxplot(doctors_per_cap~V17, data = mydata)$stats # slightly higher number of doctors in ne? 
```
```{r}
kruskal.test(doctors_per_cap~factor(V17), data = mydata) # doctors per cap not significant
```

V9: number of hospital beds 
```{r}
# test assumptions 
mydata$hospitals_per_cap = mydata$V9/mydata$V5

hist(mydata$hospitals_per_cap, breaks = 50) # right skewed
leveneTest(hospitals_per_cap~factor(V17), data = mydata) # unequal variance
```
```{r}
boxplot(V9/V5~V17, data = mydata)$stats # pretty different between regions
```
```{r}
kruskal.test(V9/V5~factor(V17), data = mydata) # number of hospital beds per capita is significant
```

V10: total serious crimes
```{r}
mydata$crimes_per_cap = mydata$V10/mydata$V5

# test assumptions 
hist(mydata$crimes_per_cap, breaks = 50) # fairly normal
leveneTest(crimes_per_cap~factor(V17), data = mydata) # unequal variance, but might be ok for anova
var(s$crimes_per_cap )
var(ne$crimes_per_cap )
var(nc$crimes_per_cap )
var(w$crimes_per_cap)
```
```{r}
boxplot(V10/V5~V17, data = mydata)$stats # very different between regions
```

```{r}
summary(aov(V10/V5~factor(V17), data = mydata)) # number of crimes per capita is significant
```
```{r}
kruskal.test(crimes_per_cap~V17, data = mydata) # pop not significant
```

V11: Percent high school graduates
```{r}
# test assumptions 
hist(mydata$V11, breaks = 50) # slightly left skewed
leveneTest(V11~factor(V17), data = mydata) # equal variance
```
```{r}
boxplot(V11~V17, data = mydata)$stats # pretty different between regions
```
```{r}
summary(aov(V11~factor(V17), data = mydata)) # high school education rate is significant
```
```{r}
kruskal.test(V11~factor(V17), data = mydata)
```

V12: percent bachelor's degree
```{r}
# test assumptions 
hist(mydata$V12, breaks = 50) # slightly right skewed
leveneTest(V12~factor(V17), data = mydata) # equal variance
```
```{r}
# could probably bin some of these 
boxplot(V12~V17, data = mydata)$stats # ne and w have high education levels
```
```{r}
summary(aov(V12~factor(V17), data = mydata)) # but college grads not significant
```
```{r}
kruskal.test(V12~factor(V17), data = mydata)
```

V13: Percent below poverty level
```{r}
# test assumptions 
hist(mydata$V13, breaks = 50) # very right skewed
leveneTest(V13~factor(V17), data = mydata) # unequal variance 
```
```{r}
boxplot(V13~V17, data = mydata)$stats # poverty level also very characteristic, could see if correlates with education 
```

```{r}
kruskal.test(V13~V17, data = mydata) # % below poverty level significant
```

V14: percent unemployment 
```{r}
# test assumptions 
hist(mydata$V14, breaks = 50) # slightly right skewed
leveneTest(V14~factor(V17), data = mydata) # unequal variance
```
```{r}
boxplot(V14~V17, data = mydata)$stats # percent unemployment almost significant

summary(aov(V14~factor(V17), data = mydata)) 
```
```{r}
kruskal.test(V14~V17, data = mydata) # unemployment is significant
```

V15: Per capita income
```{r}
# test assumptions 
hist(mydata$V15, breaks = 50) # fairly normal
leveneTest(V15~factor(V17), data = mydata) # not equal variance, but could be fine because anova is fairly robust to this 
var(s$V15)
var(ne$V15)
var(nc$V15)
var(w$V15)
```
```{r}
boxplot(V15~V17, data = mydata)$stats 

summary(aov(V15~factor(V17), data = mydata)) # per capita income is significant
```
```{r}
kruskal.test(V15~V17, data = mydata) # per capita income is significant
```

V16: total personal income 
```{r}
# test assumptions 
hist(mydata$V16, breaks = 50) # extremely right skeweed
leveneTest(V16~factor(V17), data = mydata) # equal variance
```
```{r}
boxplot(V16~V17, data = mydata)$stats
```
```{r}
summary(aov(V16~factor(V17), data = mydata)) # total personal income is significant
```

```{r}
kruskal.test(V16~V17, data = mydata) # total personal income is significant
```