---
title: "ADA Final"
author: "Sophie Bair"
date: "4/8/2021"
output: pdf_document
---

Significant
* Land area: larger in West 
* Number of hospital beds per capita: lower in west, makes sense
* Number of serious crimes per capita: higher in south, Lowest in NE 
* Percent high school graduates: lowest in south
* Percent below poverty level: highest in south and west, lowest in NE
* Per capita income: highest in NE 
* Personal income: “
Almost significant
* Percent unemployment

```{r}
data<-read.table("../data/APPENC02.txt", sep="", header=FALSE)
UNI<-2244
set.seed(UNI)
index <-sample(c(1:440))
mydata<-data[index[1:250],]
head(mydata)
```

```{r}
write.csv(mydata, "../output/ada_project_df.csv")
```

```{r}
ne <- mydata[mydata$V17 == 1, ]
nc <- mydata[mydata$V17 == 2, ]
s <- mydata[mydata$V17 == 3, ]
w <- mydata[mydata$V17 == 4, ]
nrow(ne)
nrow(nc)
nrow(s)
nrow(w)
```

```{r}
unique(w$V2) # 52 counties in new england
# 62 counties in nc
# 84 in s
# 43 in w
```
```{r}
boxplot(log(V4)~V17, data = mydata)$stats # west has large land area
```
```{r}
summary(aov(log(V4)~factor(V17), data = mydata))
```
```{r}
# test assumptions 
#hist(mydata$V4, breaks = 50) # not normal! 
hist(mydata$V4, breaks = 50) # not normal! 

```

```{r}
library(car)
leveneTest(V4~factor(V17), data = mydata) # not equal variance either, prob better to use k-w instead
```

```{r}
kruskal.test(V4~factor(V17), data = mydata) # still significant
```

```{r}
#boxplot(V5~V17, data = mydata)
boxplot(V5~V17, data = mydata)$stats # ne has large population
```
```{r}
summary(aov(V5~factor(V17), data = mydata)) # pop not significant
```


```{r}
boxplot(V6~V17, data = mydata)$stats # pretty consistent percent young
```
```{r}
summary(aov(V6~factor(V17), data = mydata)) # younger pop not significant
```

```{r}
boxplot(V7~V17, data = mydata)$stats # slightly higher older pop in ne
```
```{r}
summary(aov(V7~factor(V17), data = mydata)) # older pop not significant
```

```{r}
mydata$doctors_per_cap = mydata$V8/mydata$V5
boxplot(doctors_per_cap~V17, data = mydata)$stats # slightly higher number of doctors in ne? 
```
```{r}
summary(aov(doctors_per_cap~factor(V17), data = mydata)) # doctors per cap not significant
```

```{r}
boxplot(V9/V5~V17, data = mydata)$stats # pretty different between regions
```
```{r}
summary(aov(V9/V5~factor(V17), data = mydata)) # number of hospital beds per capita is significant
```

```{r}
boxplot(V10/V5~V17, data = mydata)$stats # very different between regions
```
```{r}
summary(aov(V10/V5~factor(V17), data = mydata)) # number of crimes per capita is significant
```
```{r}
boxplot(V11~V17, data = mydata)$stats # pretty different between regions
```
```{r}
summary(aov(V11~factor(V17), data = mydata)) # looks like south has significantly lower high school education rates 
```

```{r}
# could probably bin some of these 
boxplot(V12~V17, data = mydata)$stats # ne and w have high education levels
```
```{r}
summary(aov(V12~factor(V17), data = mydata)) # but college grads not significant
```

```{r}
boxplot(V13~V17, data = mydata)$stats # poverty level also very characteristic, could see if correlates with education 
```
```{r}
summary(aov(V13~factor(V17), data = mydata)) # older pop not significant
```


```{r}
boxplot(V14~V17, data = mydata)$stats # percent unemployment almost significant

summary(aov(V14~factor(V17), data = mydata)) 
```

```{r}
boxplot(V15~V17, data = mydata)$stats # per capita income is significant

summary(aov(V15~factor(V17), data = mydata)) 
```


```{r}
boxplot(V16~V17, data = mydata)$stats # personal income is significant

summary(aov(V15~factor(V17), data = mydata)) 
```