---
title: Exploratory Data Analysis on County Crime Dataset for Advanced Data Analysis
author: "Amir Idris"
output:
  pdf_document: default
  html_notebook: default
---
```{r}
library(dplyr)
```


First, let's load the data.

```{r}
crime <- read.table("../data/APPENC02.txt", sep="", header=FALSE)
UNI <- 2244
set.seed(UNI)
index <- sample(c(1:440))
crime <- crime[index[1:250],]

summary(crime)
```

Let's rename the columns to make it easier for ourselves
```{r}

column.names <- c("Id", "County Name", "State", "Land Area", "Total Pop.", "Percent of Pop. Age 18-34", "Percent of Pop. Age >= 65", "Num of Active Physicians", "Num of Hospital Beds", "Total Serious Crimes", "Percent High School Graduates", "Percent Bachelor's Degree", "Percent Below Poverty Level", "Percent Unemployment", "Per Capita Income", "Total personal income", "Geographic Region")

colnames(crime) <- column.names
head(crime)
```

We have quite a few variables. Let's take this step by step. First, let's plot the variables that are likely to have a wide-spread, since we may have to log-transform those later. Let's take a look at land area, total population, and total serious crimes by county.

```{r}
par(mfrow=c(3,1))

# Land Area Plot
land_df <- crime[order(crime$`Land Area`), ]

barplot(land_df$`Land Area`, names.arg = land_df$`County Name`, xlab = "County", ylab = "Land Area in Sq. Miles", main = "Land Area for each County")

# Total Population Plot
pop_df <- crime[order(crime$`Total Pop.`), ]

barplot(pop_df$`Total Pop.`, names.arg = pop_df$`County Name`, xlab = "County", ylab = "Total Population", main = "Total 1990 Population for each County")

# Total Serious Crimes
crime_df <- crime[order(crime$`Total Serious Crimes`), ]

barplot(crime_df$`Total Serious Crimes`, names.arg = crime_df$`County Name`, xlab = "County", ylab = "Total Serious Crimes", main = "Total 1990 Serious Crimes for each County")
```

As expected, all of these variables have an enormous spread spanning multiple orders of magnitude. In order to use these in regression models later, they'll have to be log-transformed.


Now, before we move on to modeling, let's take a look at how some of the variables differ by region. 
```{r}
# Select variables
selected_data <- subset(crime, select = -c(Id, `County Name`, State,`Land Area`, `Total Pop.`, `Total Serious Crimes`))
d <- dim(selected_data)[2]
par(mfrow=c(ceiling(sqrt(d)), ceiling(sqrt(d))))
#crime$`Geographic Region` <- as.factor(crime$`Geographic Region`)

for(i in 1:(d-1)){
  #title_string <- paste("Box and Whisker of ", colnames(selected_data)[i], " by Region", sep = "")
  #formula_string <- paste("Geographic Region ~ ", colnames(selected_data)[i], sep = "")
  #plot_formula <- as.formula(formula_string)
  
  boxplot(selected_data[, i] ~ `Geographic Region`, data = selected_data, main=title_string,
   xlab="Geographic Region", ylab=colnames(selected_data)[i])
}
```












