---
title: Polytomous Logistic Regression Model to Predict Region in County Crime Dataset
author: "Amir Idris"
output:
  pdf_document: default
  html_notebook: default
---
```{r}
if (!require("dplyr")){
  install.packages("dplyr")
}

if (!require("nnet")){
  install.packages("nnet")
}

library(dplyr)
library(nnet)
```


First, let's load the data.

```{r}
crime <- read.table("../data/APPENC02.txt", sep="", header=FALSE)
UNI <- 2244
set.seed(UNI)
index <- sample(c(1:440))
crime <- crime[index[1:250],]

# Rename column names to reflect meaning
column.names <- c("Id", "County Name", "State", "Land Area", "Total Pop.", "Percent of Pop. Age 18-34", "Percent of Pop. Age >= 65", "Num of Active Physicians", "Num of Hospital Beds", "Total Serious Crimes", "Percent High School Graduates", "Percent Bachelor's Degree", "Percent Below Poverty Level", "Percent Unemployment", "Per Capita Income", "Total personal income", "Geographic Region")

colnames(crime) <- column.names

# Add crimes per person as column
crimes_per_person <- crime$`Total Serious Crimes`/crime$`Total Pop.`
crime <- as.data.frame(cbind(crime, crimes_per_person))

crime$`Geographic Region` <- as.factor(crime$`Geographic Region`) #Ensure our outcome variable is treated as a factor variable
with(crime, levels(`Geographic Region`))

head(crime)
```

As found in our ANOVA tests, the following predictors have significant or near-significant differences between regions, at an alpha-level of 0.05. So, we can use these in our model.

<b> Significant </b>
* Land area: larger in West 
* Number of hospital beds per capita: lower in west, makes sense
* Number of serious crimes per capita: higher in south, Lowest in NE 
* Percent high school graduates: lowest in south
* Percent below poverty level: highest in south and west, lowest in NE
* Per capita income: highest in NE 
* Personal income: “
<b> Almost significant </b>
* Percent unemployment

Next, lets split the data into training and testing, and 80/20 split.

```{r}
# Credit to https://stackoverflow.com/questions/17200114/how-to-split-data-into-training-testing-sets-using-sample-function
train_index <- sample.int(n = nrow(crime), size = floor(.80*nrow(crime)), replace = F)
train_crime <- crime[train_index, ]
test_crime  <- crime[-train_index, ]

print(dim(train_crime))
print(dim(test_crime))
```

Now let's fit our Logistic Regression Model

```{r}
mod0 <- multinom(`Geographic Region` ~ `Land Area` + `Num of Hospital Beds` + crimes_per_person + `Percent High School Graduates` + `Percent Below Poverty Level` + `Per Capita Income` + `Percent Unemployment`, data = train_crime)
summary(mod0)
```


Let's see if we can judge the accuracy of this model on the training and testing datasets

```{r}
# Credit to https://datasciencebeginners.com/2018/12/20/multinomial-logistic-regression-using-r/

# Predicting the values for train dataset
train_crime$predicted <- predict(mod0, newdata = train_crime, "class")

# Building classification table
ctable_train <- table(train_crime$`Geographic Region`, train_crime$predicted)

# Calculating accuracy - sum of diagonal elements divided by total obs
train_acc <- round((sum(diag(ctable_train))/sum(ctable_train))*100,4)


# Predicting the values for test dataset
test_crime$predicted <- predict(mod0, newdata = test_crime, "class")

# Building classification table
ctable_test <- table(test_crime$`Geographic Region`, test_crime$predicted)

# Calculating accuracy - sum of diagonal elements divided by total obs
test_acc <- round((sum(diag(ctable_test))/sum(ctable_test))*100,4)

print(paste("Training Accuracy: ", train_acc, "%", sep=""))
print(paste("Testing Accuracy: ", test_acc, "%", sep=""))
```

While our accuracy is not incredibly high, we are performing much better than random, which with 4 categories would be 25%. Let's see which of our coefficients are significant.

```{r}
# Credit to https://stats.idre.ucla.edu/r/dae/multinomial-logistic-regression/

z <- summary(mod0)$coefficients/summary(mod0)$standard.errors
p <- (1 - pnorm(abs(z), 0, 1)) * 2 #testing if B_i = 0 vs. B_i /= 0
p
```



We can see that there are only three coefficients that are not significant at an alpha-level of 0.05: num of hospital beds for level 2, percent of high school graduates for level 3, and per capita income for level 4. Since these coefficients are in comparison to our baseline level of the Northeast, then this tells us that these features in these regions do not differ significantly from the northeast in our data according to our model.











